---
title: "Heatmaps"
author: "Bismark Appiah"
date: "14 December 2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Run Markdown file "Code_for_Electroporation_Dataset.Rmd" before running the scripts below 
# Load recquired packages
```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(ComplexHeatmap)
library(circlize)
library(Seurat)
```

# 2. Heatmap of selected top positive markers of AP, TTS, BP / TTS, BP, N
```{r}
# Load file containing selected markers "TC_ID.csv"
newgene <- as.character(read.csv2("TC_ID.csv")[,1])

# Subset seurat object of electroporated cells for AP, BP, and TTS clusters
# Select clusters to include for heatmap based on idents
kty <- subset(sce_hvg.seurat, idents = c("Apical Progenitors", "Basal Progenitors", "TTS")) 
z <- DoHeatmap(kty,features = newgene,label = F,
               group.colors = brewer.pal(8,'Dark2')) + NoLegend()
y <- z$data %>% drop_na()
x <- y %>% group_by(Identity) %>% dplyr::select(Feature, Cell,Identity , Expression) %>%
  tidyr::spread(key = Feature, value = Expression)

# Subset seurat object of electroporated cells for TTS, BP, and N clusters
kty_2 <- subset(sce_hvg.seurat, idents = c("TTS", "Basal Progenitors", "Neurons")) 
zz <- DoHeatmap(kty_2,features = newgene,label = F,
               group.colors = brewer.pal(8,'Dark2')) + NoLegend()
yy <- zz$data %>% drop_na()
xx <- yy %>% group_by(Identity) %>% dplyr::select(Feature, Cell,Identity , Expression) %>%
  tidyr::spread(key = Feature, value = Expression)

```

# 3. Generate heatmaps using ComplexHeatmap 
```{r}
# Heatmap for AP, TTS, BP
res_list = x
type = res_list$Identity
type <- sort.int(type)
res_list <- t(as.matrix(res_list[, -c(1,2)]))

set.seed(342)
ha <- HeatmapAnnotation(type = type, annotation_name_side = "left",
                      show_annotation_name = T,show_legend = T)

ht_list1 <- Heatmap(res_list, name = "TC Markers",
                  top_annotation = ha, 
                  show_column_names = T, row_title = NULL,show_row_dend = FALSE,
                  cluster_columns = F,show_column_dend = FALSE,row_names_side = "left"
                  ,border = T,row_names_gp = gpar(fontsize = 10),cluster_rows = T,column_gap = 5) 


# Heatmap for TTS, BP, N clusters
res_list_2 = xx
type_2 = res_list_2$Identity
type_2 <- sort.int(type_2)
res_list_2 <- t(as.matrix(res_list_2[, -c(1,2)]))


ha_2 <- HeatmapAnnotation(type = type_2, annotation_name_side = "left",
                      show_annotation_name = TRUE,show_legend = T)

ht_list2 <- Heatmap(res_list_2, name = "TC Markers",
                  top_annotation = ha_2, 
                  show_column_names = T, row_title = NULL,show_row_dend = FALSE,
                  cluster_columns = F,show_column_dend = FALSE,row_names_side = "left"
                  ,border = T,row_names_gp = gpar(fontsize = 10),cluster_rows = T,column_gap = 5)

```

# 4. Plot Supplementary Figure 5B, Appiah et al. | Order of columns rearranged, labels edited in inkscape for final figure in manuscript
```{r}
# Load file containing selected markers "TC_ID.csv"

draw(ht_list1, ht_gap = unit(10, "mm")) 

```

# 5. Plot Supplementary Figure 5C, Appiah et al. | Order of columns rearranged, labels edited in inkscape for final figure in manuscript
```{r}
# Load file containing selected markers "TC_ID.csv"

draw(ht_list2, ht_gap = unit(10, "mm")) 

```


