---
title: Analyses of Microinjected Cells (E14.5) with Seurat Workflow for scRNA-seq
  data
author: "Bismark Appiah"
date: "10 April 2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, message=FALSE, warning=FALSE}
library(BiocStyle)
library(readr)
library(data.table)
library(dplyr)
library(patchwork)
library(tools)
library(SingleCellExperiment)
library(HDF5Array)
library(mvoutlier)
library(scran)
library(scater)
library(SC3)
library(slingshot)
library(stringr)
#library(org.Mm.eg.db)
library(mclust)
library(RColorBrewer)
library(Seurat)
library(umap)
#library(latexpdf)
library(cowplot)
```

# 1. Data preparation  
#Loading data matrix

Convert data to "SingleCellExperiment" Class.
## Define DMSO and EPZ conditions  
```{r loading, echo=TRUE} 
AMI_ALL_NEW <- readRDS("AMI_ALL_NEW.RData")

sce <- SingleCellExperiment(list(counts=as.matrix(AMI_ALL_NEW)))
rm(AMI_ALL_NEW)
sce
DM24 <- readRDS("DM24.RData")
EP24 <- readRDS("EP24.RData")

sce$condition <- "NA"
sce$condition[which(colnames(sce) %in% DM24)] <- "DMS24"
sce$condition[which(colnames(sce) %in% EP24)] <- "EPZ24"

table(sce$condition)

```



## Add mitochondrial genes and ERCC spike-ins to sce object  

There were no ERCC spike-ins used in the experiment.

```{r spikeins, include=TRUE, echo=TRUE}
isSpike(sce, "MT") <- rownames(sce)[grep("^(mt)",rownames(sce),invert=FALSE)]
isSpike(sce, "ERCC") <- rownames(sce)[grep("^(ERCC)",rownames(sce),invert=FALSE)]
sce
```

# 2. Quality control  

#Remove genes that are not expressed  
```{r removegenes, include=TRUE, echo=TRUE}
keep_feature <- rowSums(counts(sce) > 0) > 0
sce <- sce[keep_feature,]
```

#Calculate quality metrics  
It seems like there were no ERCC spike-ins used.
```{r qualitymetrics, include=TRUE, echo=TRUE}
sce <- calculateQCMetrics(sce)
```

```{r qualplot, include=TRUE, fig.height=10, fig.width=10}
multiplot(cols=2,
    plotColData(sce, x="condition", y="total_counts"),
    plotColData(sce, x="condition", y="total_features_by_counts"),
    plotColData(sce, x="condition", y="pct_counts_MT")
)
```
Total counts and distribution of counts for cells in both conditions looks good
We can proceed with further analysis

# Remove ERCC spike-ins since they were not used.
```{r }
sce <- sce[grep("^(ERCC|Gm|Rik)",row.names(sce),invert=TRUE),]
```


# Visualize library sizes  

```{r, include=TRUE}
hist(
    sce$total_counts,
    breaks = 100
)
abline(v = 2500, col = "red")
```

# 3. Filter cells by library size > 2500  

```{r, include=TRUE}
filter_by_total_counts <- (sce$total_counts > 2500)
table(filter_by_total_counts)
```

```{r, include=TRUE}
sce <- sce[,filter_by_total_counts]
```

# 4. Visualize the total number of unique genes detected in each sample  

```{r, include=TRUE}
hist(
    sce$total_features_by_counts,
    breaks = 100
)
abline(v = 1000, col = "red")
```

# 5. Filter cells by expressed features > 1000  

```{r, include=TRUE}
filter_by_expr_features <- (sce$total_features_by_counts > 1000)
table(filter_by_expr_features)
```

```{r, include=TRUE}
sce <- sce[,filter_by_expr_features]
```

Cells with outlier values for mitochondrial genes are identified based on some number of MADs from the median value.

```{r, include=TRUE}
high.mt <- isOutlier(sce$pct_counts_MT, nmads=3, type="higher", batch=sce$condition)
data.frame(HighMito=sum(high.mt))
```

We only retain cells that pass all of the specified criteria.
Of course, this involves some assumptions about independence from biology. 
(For example, don't use the mitochondrial proportions if the number/activity of mitochondria changes between cell types.)

```{r, include=TRUE}
discard <- high.mt
data.frame(TotalLost=sum(discard), TotalLeft=sum(!discard))
```

We toss out the cells that we consider to be low-quality, and keep the rest.

```{r, include=TRUE}
sce <- sce[,!discard]
ncol(sce)
```

https://scrnaseq-course.cog.sanger.ac.uk/website/cleaning-the-expression-matrix.html#exprs-qc

```{r, include=TRUE}
#plotColData(sce, x = "total_features_by_counts", y = "pct_counts_MT")
```

# 6. Examining the genes

We inspect the distribution of log-mean counts across all genes.
The peak represents the bulk of moderately expressed genes while the rectangular component corresponds to lowly expressed genes.

```{r, include=TRUE}
ave.counts <- calcAverage(sce)
hist(log10(ave.counts), breaks=100, main="", col="grey80",
    xlab=expression(Log[10]~"average count"))
```

# 7. Show top 30 of most highly expressed genes

We also look at the identities of the most highly expressed genes.
This should generally be dominated by constitutively expressed transcripts, such as those for ribosomal or mitochondrial proteins.
The presence of other classes of features may be cause for concern if they are not consistent with expected biology.

```{r fig.height=7, fig.width=7}
plotHighestExprs(sce, n=30)
```

```{r, include=TRUE}
clusters <- quickCluster(sce,use.ranks=FALSE)
sce <- computeSumFactors(sce, cluster=clusters)
```

From: Lun et al. (2016) A step-by-step workflow for low-level analysis of single-cell RNA-seq data with Bioconductor.

"In this case, the size factors are tightly correlated with the library sizes for all cells. This suggests that the
systematic differences between cells are primarily driven by differences in capture efficiency or sequencing depth. Any
DE between cells would yield a non-linear trend between the total count and size factor, and/or increased scatter around
the trend. This does not occur here as strong DE is unlikely to exist within a homogeneous population of cells."

In our case there is also a linear trend, but with some scatter around it.

```{r, include=TRUE}
plot(sizeFactors(sce), sce$total_counts/1e3, log="xy",
ylab="Library size (thousands)", xlab="Size factor")
```

```{r, include=TRUE}
sce <- scater::normalize(sce)
sce
```


# 8. Modelling the technical component of variation

We identify HVGs to focus on the genes that are driving heterogeneity across the population of cells.
This requires estimation of the variance in expression for each gene, followed by decomposition of the variance into biological and technical components.

From: Lun et al. (2016) A step-by-step workflow for low-level analysis of single-cell RNA-seq data with Bioconductor.

"HVGs  are  defined  as  genes  with  biological  components  that  are  significantly  greater  than  zero  at  a  false  discovery rate (FDR) of 5%. These genes are interesting as they drive differences in the expression profiles between cells, and should be prioritized for further investigation. In addition, we only consider a gene to be a HVG if it has a biological component greater than or equal to 0.5. For transformed expression values on the log2 scale, this means that the average  difference  in  true  expression  between  any  two  cells  will  be  at  least  2-fold.  (This  reasoning  assumes  that the true log-expression values are Normally distributed with variance of 0.5. The root-mean-square of the difference between two values is treated as the average log2-fold change between cells and is equal to unity.) We rank the results by the biological component to focus on genes with larger biological variability."

```{r, include=TRUE}
var.fit <- trendVar(sce, method="loess", loess.args=list(span=0.05), use.spikes=FALSE)
var.out <- decomposeVar(sce, var.fit)
#head(var.out)    
```

"We can have a look at the fitted trend.
Some tinkering may be required to get a good fit, usually by modifying `span=`."

```{r hvgplothsc, include=TRUE}
plot(var.out$mean, var.out$total, pch=16, cex=0.6, xlab="Mean log-expression", 
    ylab="Variance of log-expression")
curve(var.fit$trend(x), add=TRUE, col="dodgerblue", lwd=2)
cur.spike <- isSpike(sce)
points(var.out$mean[cur.spike], var.out$total[cur.spike], col="red", pch=16)
```

# Extracting highly variable genes (HVG)
I selected 800 of the HVGs extracted for clustering. This helps to focus on the genes that are driving heterogeneity across the population of cells.
```{r, include=TRUE}
hvg.out <- var.out[with(var.out, order(bio,FDR, decreasing = TRUE)),][1:800,]
hvg.out <- hvg.out[order(hvg.out$bio, decreasing = TRUE),]
nrow(hvg.out)
```


# 9. Checking the distribution of expression values for the top HVGs
This ensures that the variance estimate is not being dominated by one or two outlier cells.

```{r fig.height=5, fig.width=7, include=TRUE}
plotExpression(sce, rownames(hvg.out)[1:10])
```

# 10. Extract highly variable genes (HVG)  
 
```{r, include=TRUE}
sce_hvg <- sce[row.names(hvg.out),]
```

# 11. Seurat clustering analysis  

Convert SingleCellExperiment object into Seurat object.

```{r}
rownames(sce_hvg) <- str_remove(rownames(sce_hvg), "[|]")
```


```{r}
sce_hvg.seurat <- as.Seurat(x = sce_hvg)
```

Scaling data before running PCA.

```{r}
all.genes <- rownames(sce_hvg.seurat)
sce_hvg.seurat <- ScaleData(sce_hvg.seurat, features = all.genes)
```


```{r removeercc}
sce_hvg.seurat <- sce_hvg.seurat[grep("^(mt)",row.names(sce_hvg.seurat),invert=TRUE),]
```


# 12. Run PCA on Seurat object

```{r}
sce_hvg.seurat <- RunPCA(sce_hvg.seurat, npcs = 60, features = rownames(sce_hvg.seurat))
```

# 13. Inspecting first two principal components

```{r}
VizDimLoadings(sce_hvg.seurat, dims = 1:2, reduction = "pca")
```

```{r include=FALSE}
DimPlot(sce_hvg.seurat, reduction = "pca")
```

# 14. Showing heatmap of first 10 principal components

```{r}
DimHeatmap(sce_hvg.seurat, dims = 1:10, cells = 500, balanced = TRUE)
```

Heatmap allows to see some structure in the data. It suggest that the first 8 PCs include some structure. These can then be used for the cluster analysis.

# 15. Determine the dimensionality of the dataset  

This part is based on: https://satijalab.org/seurat/v3.0/pbmc3k_tutorial.html

"To overcome the extensive technical noise in any single feature for scRNA-seq data, Seurat clusters cells based on their PCA scores, with each PC essentially representing a ‘metafeature’ that combines information across a correlated feature set. The top principal components therefore represent a robust compression of the dataset. However, how many components should we choose to include? 10? 20? 100?

In Macosko et al, we implemented a resampling test inspired by the JackStraw procedure. We randomly permute a subset of the data (1% by default) and rerun PCA, constructing a ‘null distribution’ of feature scores, and repeat this procedure. We identify ‘significant’ PCs as those who have a strong enrichment of low p-value features."

```{r}
sce_hvg.seurat <- JackStraw(sce_hvg.seurat, num.replicate = 100, dims = 20)
sce_hvg.seurat <- ScoreJackStraw(sce_hvg.seurat, dims = 1:20)
```

```{r}
JackStrawPlot(sce_hvg.seurat, dims = 1:20, xmax = 0.75, ymax = 1.0)
```

The drop of the p-value at PC8 suggest to use the first 8 dimensions for clustering.

```{r include=FALSE}
ElbowPlot(sce_hvg.seurat, ndims = 15, reduction = "pca")
```
The more approximate elbow plot also suggests to use 8 PCs.



Cluster evaluation  

Run Seurat clustering with varying resolution parameters to assess cluster stability. Parameter is varied between 0.4 to 1.2 in steps of 0.1. The resulting clusterings are then compared to each other using the adjusted rand index (ARI).

Compute average silhouette width for resolution parameter.




# 16. Cluster the cells  

Using a resolution of 0.4 seems to be the most stable version according to the average silhouette and ARI.

```{r}
sce_hvg.seurat <- FindNeighbors(sce_hvg.seurat, dims = 1:8)
sce_hvg.seurat <- FindClusters(sce_hvg.seurat, resolution = 0.4, algorithm = 1)
```

```{r}
head(Idents(sce_hvg.seurat), 5)
```


# 17. Run UMAP for visualization in two dimensions 

```{r}
sce_hvg.seurat <- RunUMAP(sce_hvg.seurat, dims = 1:8)
DimPlot(sce_hvg.seurat, reduction = "umap")
```

# 18. Finding differentially expressed features (cluster biomarkers)  

## Markers for Cluster 0  

```{r}
# find all markers of cluster 0
cluster0.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 0)
head(cluster0.markers, n = 30)
```

# UMAP visualization of top 15 markers for Cluster 0
```{r fig.height=10, fig.width=15}
FeaturePlot(sce_hvg.seurat, features = c(row.names(cluster0.markers)[1:15]))
```


## Markers for Cluster 1

```{r}
# find all markers of cluster 1
cluster1.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 1)
head(cluster1.markers, n = 30)
```
# UMAP visualization of top 10 markers for Cluster 1
```{r fig.height=10, fig.width=15}
FeaturePlot(sce_hvg.seurat, features = c(row.names(cluster1.markers)[1:15]))
```

## Markers for Cluster 2  

```{r}
# find all markers of cluster 2
cluster2.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 2)

#write.csv2(cluster2.markers,file = "cluster2markersall.csv")
#head(cluster2.markers, n = 30)
```

# UMAP visualization of top 15 markers for Cluster 2
```{r fig.height=10, fig.width=15}
FeaturePlot(sce_hvg.seurat, features = c(row.names(cluster2.markers)[1:15]))
```

## Markers for Cluster 3  

```{r}
# find all markers of cluster 3
cluster3.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 3)
head(cluster3.markers, n = 30)
```

# UMAP visualization of top 10 markers for Cluster 3
```{r fig.height=10, fig.width=15}
FeaturePlot(sce_hvg.seurat, features = c(row.names(cluster3.markers)[17:32]))
```

## Markers for Cluster 4

```{r}
# find all markers of cluster 4
cluster4.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 4)
head(cluster4.markers, n = 30)
```
# UMAP visualization of top 15 markers for Cluster 4

```{r fig.height=10, fig.width=15}
FeaturePlot(sce_hvg.seurat, features = c(row.names(cluster4.markers)[1:15]))
```



# 19. Computing differential markers for each cluster compared to all other cells
Idea is to look at markers highly specific for each cluster. You will find cluster labels in column 6 with corresponding cluster-specific marker gene in column 7 of table below


```{r}
# find markers for every cluster compared to all remaining cells, report only the positive ones
sce_hvg.seurat.markers <- FindAllMarkers(sce_hvg.seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
sce_hvg.seurat.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)
```

# 20. Showing top 10 marker genes in each cluster as a heatmap
```{r fig.height=10, fig.width=15}
top10 <- sce_hvg.seurat.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
DoHeatmap(sce_hvg.seurat, features = top10$gene) + NoLegend()
#FeaturePlot(sce_hvg.seurat, features = "Stmn1")
```

# 21. UMAP representation of cells in each condition (DMSO/EPZ)
From here, we can compute proportions of cells for each cluster.
```{r fig.height=5, fig.width=10}
DimPlot(sce_hvg.seurat, reduction = "umap", split.by = "condition", pt.size = 1.5)
```




# 22. Using Slingshot for pseudotime analysis  

From: https://www.bioconductor.org/packages/release/bioc/vignettes/slingshot/inst/doc/vignette.html#using-slingshot

```{r}
library(slingshot)
sce_hvg.slingshot <- as.SingleCellExperiment(sce_hvg.seurat)
```

## Gaussian mixture model for clustering  

```{r include=FALSE}
cl1 <- Mclust(reducedDim(sce_hvg.slingshot)[,1:2])$classification
cl1 <- Idents(sce_hvg.seurat)
colData(sce_hvg.slingshot)$GMM <- cl1

plot(reducedDim(sce_hvg.slingshot)[,1:2], col = brewer.pal(9,"Set1")[cl1], pch=16, asp = 1)
```



```{r include=FALSE}
sce_hvg.slingshot <- slingshot::slingshot(sce_hvg.slingshot, clusterLabels = 'GMM', reducedDim = 'UMAP')
```

```{r}
saveRDS(file="sce_hvg.slingshot", sce_hvg.slingshot)
```

```{r}
sce_hvg.slingshot <- readRDS(file="sce_hvg.slingshot")
```

```{r}
summary(sce_hvg.slingshot$slingPseudotime_1)
```



# 23. Lineage structure visualized cluster-wise
```{r fig.height=10, fig.width=10}
plot(reducedDims(sce_hvg.slingshot)$UMAP, col = brewer.pal(9,'Set1')[sce_hvg.slingshot$seurat_clusters], pch=16, asp = 1)
lines(SlingshotDataSet(sce_hvg.slingshot), lwd=2, type = 'lineages', col = 'black')
```



```{r fig.height=10, fig.width=10}
#Computing DEGs between EPZ and DMSO
colnames(sce_hvg.seurat[[]])
Idents(object = sce_hvg.seurat) <- 'condition'
levels(sce_hvg.seurat)
#alldegs <- FindMarkers(sce_hvg.seurat, ident.1 = "EPZ24", ident.2 = "DMS24",test.use = "negbinom",verbose = T)
ordered.degs_all <- alldegs[which(alldegs$p_val_adj < 0.05),]
#write.csv2(ordered.degs_all,file = "Final_DEGS_all_INJECTED_Seurat.csv")
# Extracting seurat object for DMSO cells
sce_hvg.seurat_DMSO <- subset(sce_hvg.seurat,subset= condition=="DMS24")
#Extracting seurat object for EPZ cells
sce_hvg.seurat_EPZ <- subset(sce_hvg.seurat,subset= condition=="EPZ24")


#write.csv2(alldegs,file = "DEGS_all_INJECTED_Seurat.csv")

#Computing DEGs between conditions for each cluster

colnames(sce_hvg.seurat[[]])
Idents(object = sce_hvg.seurat) <- 'seurat_clusters'
levels(sce_hvg.seurat)
sce_hvg.seurat$condition <- paste(Idents(sce_hvg.seurat), sce_hvg.seurat$condition, sep = "_")
Idents(sce_hvg.seurat) <- "condition"

#fINDING IDENTITY OF CLUSTER 3
c3FIND <- FindMarkers(sce_hvg.seurat, ident.1 = "2_DMS24", ident.2 = "3_DMS24",test.use = "negbinom")
ordered.degs_C4 <- DEGc4[which(DEGc4$p_val_adj < 0.05),]
write.csv2(ordered.degs_C4,file = "Final_DEGS_Clust4_INJECTED_Seurat.csv")



DEGc4 <- FindMarkers(sce_hvg.seurat, ident.1 = "4_EPZ24", ident.2 = "4_DMS24",test.use = "negbinom")
ordered.degs_C4 <- DEGc4[which(DEGc4$p_val_adj < 0.05),]
write.csv2(ordered.degs_C4,file = "Final_DEGS_Clust4_INJECTED_Seurat.csv")


#write.csv2(DEGc2,file = "DEGS_c2_crazy_INJECTED_Seurat.csv")
##Violin plots of DEGssig
#VlnPlot(sce_hvg.seurat, c("Fabp7","Tubb3"))

#Assigning new cluster IDs
#current.cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7)
#new.cluster.ids <- c("CD4 T cells", "CD14+ Monocytes", "B cells", "CD8 T cells", "FCGR3A+ Monocytes", "NK cells", "Dendritic cells", "Megakaryocytes")
#pbmc@ident <- plyr::mapvalues(pbmc@ident, from = current.cluster.ids, to = new.cluster.ids)
#TSNEPlot(pbmc, do.label = T, pt.size = 0.5)


ClustXvsAP <- FindMarkers(sce_hvg.seurat, ident.1=2, ident.2=3)

ClustXvsNeuron2 <- FindMarkers(sce_hvg.seurat, ident.1=2, ident.2=1)
Increased_ClustXvsNeuron2 <- ClustXvsNeuron2[which(ClustXvsNeuron2$avg_logFC > 0),]
Decreased_ClustXvsNeuron2 <- ClustXvsNeuron2[which(ClustXvsNeuron2$avg_logFC < 0),]

write.csv2(Increased_ClustXvsNeuron2,file = "IncreasedinClustX_relative_to_Neuron2_Injected.csv")
write.csv2(Decreased_ClustXvsNeuron2,file = "DecreasedinClustX_relative_to_Neuron2_Injected.csv")



#kkkkkk <- IP.markers[which(abs(IP.markers$avg_logFC) > 1 & IP.markers$p_val_adj < 0.05),]

Increased_ClustXvsAP <- ClustXvsAP[which(ClustXvsAP$avg_logFC > 0),]
Decreased_ClustXvsAP <- ClustXvsAP[which(ClustXvsAP$avg_logFC < 0),]

#Finding cluster Markers

# find all markers of cluster 
cluster0.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 0)
ordered.markers_C0 <- cluster0.markers[which(cluster0.markers$p_val_adj < 0.05),]
#write.csv2(ordered.markers_C4,file = "Markers_Clust4_INJECTED.csv")

ordered.markers_C0_DOWN <- ordered.markers_C0[which(ordered.markers_C0$avg_logFC < 0),]
write.csv2(ordered.markers_C0_DOWN,file = "Markers_Clust0_INJECTED_DOWN.csv")

ordered.markers_C0_UP <- ordered.markers_C0[which(ordered.markers_C0$avg_logFC > 0),]
write.csv2(ordered.markers_C0_UP,file = "Markers_Clust0_INJECTED_UP.csv")


# find all markers of cluster 
cluster1.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 1)
ordered.markers_C1 <- cluster1.markers[which(cluster1.markers$p_val_adj < 0.05),]
#write.csv2(ordered.markers_C4,file = "Markers_Clust4_INJECTED.csv")

ordered.markers_C1_DOWN <- ordered.markers_C1[which(ordered.markers_C1$avg_logFC < 0),]
write.csv2(ordered.markers_C1_DOWN,file = "Markers_Clust1_INJECTED_DOWN.csv")

ordered.markers_C1_UP <- ordered.markers_C1[which(ordered.markers_C1$avg_logFC > 0),]
write.csv2(ordered.markers_C1_UP,file = "Markers_Clust1_INJECTED_UP.csv")





# find all markers of cluster 
cluster2.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 2)
ordered.markers_C2 <- cluster2.markers[which(cluster2.markers$p_val_adj < 0.05),]
#write.csv2(ordered.markers_C4,file = "Markers_Clust4_INJECTED.csv")

ordered.markers_C2_DOWN <- ordered.markers_C2[which(ordered.markers_C2$avg_logFC < 0),]
write.csv2(ordered.markers_C2_DOWN,file = "Markers_Clust2_INJECTED_DOWN.csv")

ordered.markers_C2_UP <- ordered.markers_C2[which(ordered.markers_C2$avg_logFC > 0),]
write.csv2(ordered.markers_C2_UP,file = "Markers_Clust2_INJECTED_UP.csv")





# find all markers of cluster 
cluster3.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 3)
ordered.markers_C3 <- cluster3.markers[which(cluster3.markers$p_val_adj < 0.05),]
#write.csv2(ordered.markers_C4,file = "Markers_Clust4_INJECTED.csv")

ordered.markers_C3_DOWN <- ordered.markers_C3[which(ordered.markers_C3$avg_logFC < 0),]
write.csv2(ordered.markers_C3_DOWN,file = "Markers_Clust3_INJECTED_DOWN.csv")

ordered.markers_C3_UP <- ordered.markers_C3[which(ordered.markers_C3$avg_logFC > 0),]
write.csv2(ordered.markers_C3_UP,file = "Markers_Clust3_INJECTED_UP.csv")





# find all markers of cluster 
cluster4.markers <- FindMarkers(sce_hvg.seurat, ident.1 = 4)
ordered.markers_C4 <- cluster4.markers[which(cluster4.markers$p_val_adj < 0.05),]
#write.csv2(ordered.markers_C4,file = "Markers_Clust4_INJECTED.csv")

ordered.markers_C4_DOWN <- ordered.markers_C4[which(ordered.markers_C4$avg_logFC < 0),]
write.csv2(ordered.markers_C4_DOWN,file = "Markers_Clust4_INJECTED_DOWN.csv")

ordered.markers_C4_UP <- ordered.markers_C4[which(ordered.markers_C4$avg_logFC > 0),]
write.csv2(ordered.markers_C4_UP,file = "Markers_Clust4_INJECTED_UP.csv")



FeaturePlot(sce_hvg.seurat, features = c("Tubb3","Tbr1"))

DotPlot(sce_hvg.seurat, features = c("Fabp7","Sox2","Insm1","Eomes","Tubb3","Tbr1","Bcl11b","Satb2")) + RotatedAxis()

```



# Computing contribution of each condition to proportion of cells in each cluster
```{r echo=TRUE}
clust <- sce_hvg.seurat$seurat_clusters
clust_pts <- data.frame(names=names(clust),cluster=clust)
clust_pts$condition <- substr(clust_pts$names, start = 0, stop = 6)
#levels(clust_pts$cluster)
table_clust_pts <- table(clust_pts$condition, clust_pts$cluster)

```





# Renaming Clusters
```{r echo=TRUE}

colnames(sce_hvg.seurat[[]])
Idents(object = sce_hvg.seurat) <- 'seurat_clusters'
levels(sce_hvg.seurat)

# Extracting seurat object for DMSO cells
sce_hvg.seurat_DMSO <- subset(sce_hvg.seurat,subset= condition=="DMS24")
#Extracting seurat object for EPZ cells
sce_hvg.seurat_EPZ <- subset(sce_hvg.seurat,subset= condition=="EPZ24")

#New Cluster Names
colnames(sce_hvg.seurat[[]])
Idents(object = sce_hvg.seurat) <- 'seurat_clusters'
levels(sce_hvg.seurat)

current.cluster.ids <- c(0, 1, 2, 3, 4)
new.cluster.ids <- c("Intermediate progenitors","Neurons II","Transit cells", "Apical progenitors","Neurons I")
names(new.cluster.ids) <- levels(sce_hvg.seurat)
sce_hvg.seurat <- RenameIdents(sce_hvg.seurat, new.cluster.ids)
#choose a color pallette for clusters
t <- brewer.pal(n = 6,name = "Dark2")
DimPlot(sce_hvg.seurat, reduction = "umap", label = TRUE, cols = t,pt.size = 2.5,label.size = 8) + NoLegend()


###########
#current.cluster.ids <- c(0, 1, 2, 3, 4)
#new.cluster.ids <- c("Neuron 1 (Cluster 4)", "Neuron 2 (Cluster 5)", "Unknown (Cluster 3)", "Apical Progenitor (Cluster 1)", "Intermediate Progenitor (Cluster 2)")
#sce_hvg.seurat@active.ident <- plyr::mapvalues(sce_hvg.seurat@active.ident, from = current.cluster.ids, to = new.cluster.ids)


#genestosee <- c("Sox9","Top2a","Mki67","Aldoc","Slc1a3","Ckap2l","Neurog2","Eomes","Cadm1","Rnd2")


Sox9
Top2a
Mki67
Ckap2l
Aldoc
Slc1a3
Ckap2l
Neurog2
Eomes


Cadm1
Rnd2



DotPlot(sce_hvg.seurat, features = genestosee) + RotatedAxis()




DotPlot(sce_hvg.seurat, features = c("Hes1","Hes5","Sox9","Ptprz1","Ttyh1","Fabp7","Sox2","Nusap1","Insm1","Sox5","Eomes","Neurog2","Dcx","Neurod6","Dpysl3","Sox11","Elavl3","Tubb3","Tbr1","Bcl11b","Satb2")) + RotatedAxis()

RidgePlot(sce_hvg.seurat, features = c("Fabp7","Sox2","Insm1","Eomes","Tubb3","Tbr1","Bcl11b","Pou3f2"), ncol = 2)


DimPlot(sce_hvg.seurat, reduction = "umap")

#alldegs <- FindMarkers(sce_hvg.seurat, ident.1 = "EPZ", ident.2 = "DMSO")
#write.csv2(alldegs,file = "DEGS_all_MICROINJECTED_Seurat.csv")
```



Ttyh1
Hes1
Sox9
Hes5
Sox2
Ptprz1
Insm1
Eomes
Rrm2
Mki67
Hmgb2
Neurog2
Elavl2
Top2a
Aldoa
Nr2f1
Gas5
Pgam1
Fgfr3
Meis2
Stmn1
Sox11
Zic1
Zic3
Zic4
Dlx2
Tubb3
Dpysl3
Elavl3
Dcx
Nhlh1
Dcx
Slc17a6
Tbr1
Bcl11b
Neurod1
Tagln3
Neurod6
Neurod2






# Session info  

```{r, echo=TRUE}
sessionInfo()
```








